---
title: 'Crime In Baltimore'
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    df_print: paged
    fig_caption: true
    toc: true
    #fig_width: 7
    #fig_height: 4.5
    theme: cosmo
    highlight: tango
    #code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# - Team Members
* Oytun Deniz
* Selen Korkut Ozer
* Samet Boyaci
* Yigitcan Tunay

# - Introduction

Violent crime in the United States has fallen over the last two decades. Even though the violent crime rate has been decreasing since 1990, the United States top the ranking of countries with the most prisoners. And Baltimore, Maryland is infamous for its very high crime rates. According to Forbes, Baltimore is the seventh dangerous city in the cities of United States. The violent crime rate is 1417 per 100.000 residents in a population of 626.848 people in 2017 so Baltimore is the one of the cities which is carrying high risk for local people. That is why, we thought that it would be interesting to analyze crimes in Baltimore.

# - Project Details

Something

# - Preparation

## - Loading Data

```{r}
setwd('~/gitProjects/DAE/') #Oytun
bcd <- read.csv('Baltimore_Crime_Data.csv') #Oytun

#getwd()
#setwd('C:/Users/Sancak/Documents/R')
#bcd <- read.csv('Baltimore_Crime_Data.csv', header = T,  sep=",")
str(bcd)
```

## - Loading Libraries

```{r}
library("data.table")
library("ggplot2")
library("dplyr")
library("ggplot2")
library("ggmap")
library("maps")
library("mapdata")
library("sqldf")
```

## - Reformating features

### - Dates
```{r } 
bcd$CrimeDate<-as.Date(bcd$CrimeDate,format="%m/%d/%Y")
bcd$CrimeTimehour<- as.integer(substr(bcd$CrimeTime,0,2))
bcd$Year <- as.numeric (format(bcd$CrimeDate,"%Y"))
bcd$Month=as.numeric (format(bcd$CrimeDate,"%m"))
bcd$Day=as.numeric (format(bcd$CrimeDate,"%d"))
#Update 00:00:00 times as 0,6 - Oytun
bcd$CrimeTimeLine[bcd$CrimeTimehour == 0] <-  '(0,6]'
```

### - Inside.OutSide
* The blank outside.inside rows are set as NA
* Inside.Outside column has 4 unique rows. O, I, Outside, Inside. We standardized these rows as O -> Outside, I -> Inside

```{r}
#Oytun
bcd$Inside.Outside[bcd$Inside.Outside=="O"] <- "Outside"
bcd$Inside.Outside[bcd$Inside.Outside=="I"] <- "Inside"
bcd$Inside.Outside[bcd$Inside.Outside==""] <- NA
```

### - Location
* The blank location rows are set as NA

```{r}
bcd$Location[bcd$Location==""] <- NA
```

### - District

* The blank district rows are set as NA

```{r }
#Samet
bcd$District[bcd$District == ""] <- NA

#District Feature Reordered
bcd$District <- factor(bcd$District,levels = c("CENTER", "WESTERN", "NORTHWESTERN", "NORTHERN", "NORTHEASTERN", "EASTERN","SOUTHEASTERN","SOUTHERN","SOUTHWESTERN"))

```



## - New Features
We created 2 features

* Crime Group
* Crime Score
* Crime TimeLine

### - Crime Group

```{r }
#Samet
bcd$CrimeGroup[like(bcd$Description,"ROBBERY")] <- "ROBBERY"
bcd$CrimeGroup[like(bcd$Description,"ASSAULT")] <- "ASSAULT"
bcd$CrimeGroup[like(bcd$Description,"LARCENY")] <- "LARCENY"
bcd$CrimeGroup[like(bcd$Description,"ARSON")] <- "ARSON"
bcd$CrimeGroup[like(bcd$Description,"RAPE")] <- "RAPE"
bcd$CrimeGroup[like(bcd$Description,"SHOOTING")] <- "SHOOTING"
bcd$CrimeGroup[like(bcd$Description,"AUTO THEFT")] <- "AUTO THEFT"
bcd$CrimeGroup[like(bcd$Description,"HOMICIDE")] <- "HOMICIDE"
bcd$CrimeGroup[like(bcd$Description,"BURGLARY")] <- "BURGLARY"
```

### - Crime Score

Crime score range is 1 to 10

```{r}
#Samet
bcd$CrimeScore[bcd$Description=="AUTO THEFT"] <-4
bcd$CrimeScore[bcd$Description=="SHOOTING"] <-8
bcd$CrimeScore[bcd$Description=="AGG. ASSAULT"] <-6
bcd$CrimeScore[bcd$Description=="COMMON ASSAULT"] <-4
bcd$CrimeScore[bcd$Description=="BURGLARY"] <-10
bcd$CrimeScore[bcd$Description=="HOMICIDE"] <-10
bcd$CrimeScore[bcd$Description=="ROBBERY - STREET"] <-4
bcd$CrimeScore[bcd$Description=="ROBBERY - COMMERCIAL"] <-5
bcd$CrimeScore[bcd$Description=="LARCENY"] <-2
bcd$CrimeScore[bcd$Description=="LARCENY FROM AUTO"] <-1
bcd$CrimeScore[bcd$Description=="ARSON"] <-6
bcd$CrimeScore[bcd$Description=="ROBBERY - CARJACKING"] <-3
bcd$CrimeScore[bcd$Description=="ASSAULT BY THREAT"] <-1
bcd$CrimeScore[bcd$Description=="RAPE"] <-7
bcd$CrimeScore[bcd$Description=="ROBBERY - RESIDENCE"] <-5 # oyt duz robbery ekledim nullar kalmis


bcd$CrimeScore <- as.factor(bcd$CrimeScore)
```

### Crime Time Line

We created 4 level from Crime Time Hour feature

00:00 - 06:00, 06:00 - 12:00, 12:00 - 18:00, 18:00 - 24:00

```{r}
bcd$CrimeTimeLine <- cut(bcd$CrimeTimehour, breaks=c(0,6,12,18,24), right=TRUE)

```

### Weapon Type

We investigated, are weapon column updatable for "null"space" values:

```{r }
#Selen
sqldf("select CrimeGroup,Weapon, count(*) from bcd group by CrimeGroup,Weapon")

bcd$Weapon[bcd$Weapon =='' & bcd$CrimeGroup =='RAPE']<- 'OTHER' #updating weapon type for rape' s

#just rape crime type were updated. The other space values in "Weapon" Column are updated "NA". For doing this, a new level to factor.

bcd$Weapon[bcd$Weapon==""] <- NA

ggplot(data=bcd , aes(Weapon)) + 
  geom_bar(aes(y = (..count..)/sum(..count..)), fill='Blue', color='Black') + 
  labs(x= "Weapon Type", y="Number of Incident") 

# The Result, "Weapon" column has many "Not Known" value (over 60% of data) The other values are not enough for a clear analysis. 

```


## Data Virsulation

### District

```{r}
ggplot(subset(bcd,!is.na(District)))+
  aes(x=District)+
  geom_bar(stat = "count",fill='red') + 
  geom_text(stat="count",aes(label=..count..),vjust=-1)+
  labs(title="Frequency of Incidents by Disctrict",x="Districts",y="Number of Incidents")


ggplot(subset(bcd,!is.na(District)  & Year!=2017))+
  aes(x=Year, color=District)+
  geom_line(stat="count")+
  scale_x_continuous(breaks = seq(2012,2017,1))+
  scale_y_continuous(breaks = seq(5000,50000,5000))+
 # scale_fill_brewer(type = 'div')+
  labs(title="Frequency of Incidents by Disctrict",x="Districts",y="Number of Incidents")
```

### Neighborhood

```{r}
bcd.arrange.Neighborhood <- bcd %>%
    group_by(Neighborhood,Year) %>%
    summarise(count=n()) %>%
  arrange(desc(count))
```


```{r}
#Especillay Downtown has highest number of Incidents
ggplot(subset(bcd.arrange.Neighborhood,Neighborhood=="Downtown" & Year!=2017))+
  aes(x=Year,y=count)+
  geom_line()  +
  geom_point()+
  scale_y_continuous(breaks = seq(1400,2000,100))+
  scale_fill_brewer(type = 'div')+
  labs(title="Number of Incidents in Downtown by Year",x="Year",y="Number of Incidents")
```

### Inside - Outside And Crime Group

```{r}
#Oytun
ggplot(data =bcd[!is.na(bcd$Inside.Outside),] , aes(x = Inside.Outside, fill= CrimeGroup))+
  geom_bar(position = 'dodge') +
  labs(title = "Number of incidents by Crime Group b\n", x = "Inside or Outside", y = "Number", color = "Crime Group\n")

```

### Inside - Outside and Crime Score

```{r}
#Oytun
ggplot(data =bcd[!is.na(bcd$Inside.Outside),] , aes(x = CrimeScore, fill = Inside.Outside))+
  geom_bar(position = 'dodge') +
  labs(title = "Number of incidents by Inside Outside b\n", x = "Crime Score", y = "Number", color = "Inside or Outside\n")

```

### Crime Score and Location

```{r }
#Oytun
bcd.District.CrimeScore <- bcd[!is.na(bcd$District),] %>% 
  group_by(District, Year) %>% 
  summarise(mean_CrimeScore = mean(as.integer(CrimeScore)))

colnames(bcd.District.CrimeScore)[1] <- "District"



ggplot(data = bcd.District.CrimeScore, aes(x = Year, y =mean_CrimeScore, color = District ) )+
  geom_line() + 
  labs(title = "Mean of Crime Score by District\n", x = "Year", y = "Average Crime Score", color = "Inside or Outside\n")

```



## Data Visualization on Map

```{r}
states <- map_data("state")


ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + 
  coord_fixed(1.3) +
  guides(fill=FALSE)

maryland <- subset(states, region == c("maryland"))

ggplot(data = maryland) + 
  geom_polygon(aes(x = long, y = lat), fill = "palegreen", color = "black") 


maryland2 <- ggplot(data = maryland, mapping = aes(x = long, y = lat, group = group)) + coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")

maryland2

sqldf("select min(Longitude),max(Longitude) from bcd")
sqldf("select min(Latitude),max(Latitude) from bcd")


counties <- map_data("county")

ca_county <- subset(counties, region == "maryland")

maryland3<- maryland2 + theme_nothing() + 
  geom_polygon(data = ca_county, fill = NA, color = "white") +
  geom_polygon(color = "black", fill = NA)


# baltimore map

baltimore_map <- maryland3 + 
  coord_fixed(xlim = c(-76.71162, -76.5285),  ylim = c(39.20041, 39.37293) )

baltimore_map

baltimore_map +
  geom_point(data=bcd,aes(x = Longitude, y = Latitude,group=Total.Incidents),colour = 'Red',
             fill="Red",pch=21, size=1, alpha=I(0.005))


baltimore_map +
  geom_point(data=bcd,aes(x = Longitude, y = Latitude,group=Total.Incidents, 
                          colour=CrimeGroup),pch=21, size=1, alpha=I(0.5)) 

```






